---
title: "Clustering YSK"
output: html_notebook
---

```{r}
library("rjson")
library(dplyr)
library(tidyverse)
# Give the input file name to the function.
MeasurementData_2209 <- fromJSON(file ="C:/Users/iced/Documents/HY/Year1/DS project/data/2209.json") %>% as.data.frame #fromJSON()
MeasurementData_2209<-gather(MeasurementData_2209,"Timestamp","Value",3:26110)
```
Remove metrics word
```{r}
MeasurementData_2209$Timestamp<-substring(MeasurementData_2209$Timestamp, 9, nchar(MeasurementData_2209$Timestamp))
MeasurementData_2209$Timestamp<-as.POSIXct(as.numeric(as.character(MeasurementData_2209$Timestamp)),origin="1970-01-01",tz="GMT")
```
Divide data into CO2 and temp
```{r}
MeasurementData_2209_co2<-MeasurementData_2209[substring(MeasurementData_2209$fields,10,12)=='co2',]
MeasurementData_2209_temp<-MeasurementData_2209[!substring(MeasurementData_2209$fields,10,12)=='co2',]
MeasurementData_2209_temp<-MeasurementData_2209_temp[!substring(MeasurementData_2209_temp$fields,15,16)=='se',]
```
distinct check
```{r}
unique(MeasurementData_2209_temp$fields)
```
Adding more time columns (hour,weekday,month)
```{r}
library(chron)
MeasurementData_2209_co2$Hour=hours(MeasurementData_2209_co2$Timestamp)
MeasurementData_2209_co2$Weekday=weekdays(MeasurementData_2209_co2$Timestamp)
MeasurementData_2209_co2$Month=months(MeasurementData_2209_co2$Timestamp)
MeasurementData_2209_co2$Room<-substring(MeasurementData_2209_co2$fields,6,8)
#same for temperature
MeasurementData_2209_temp$Hour=hours(MeasurementData_2209_temp$Timestamp)
MeasurementData_2209_temp$Weekday=weekdays(MeasurementData_2209_temp$Timestamp)
MeasurementData_2209_temp$Month=months(MeasurementData_2209_temp$Timestamp)
MeasurementData_2209_temp$Room<-substring(MeasurementData_2209_temp$fields,6,8)                                                                                                               
```
Subsetting data to be from 9 AM to 18 PM and from Monday to Friday
```{r}
MeasurementData_2209_co2 <- MeasurementData_2209_co2[ which(MeasurementData_2209_co2$Hour>=9
& MeasurementData_2209_co2$Hour <= 18), ] #ok
MeasurementData_2209_co2[which(MeasurementData_2209_co2$Weekday != "Saturday" & MeasurementData_2209_co2$Weekday != "Sunday"), ]
MeasurementData_2209_temp <- MeasurementData_2209_temp[ which(MeasurementData_2209_temp$Hour>=8
& MeasurementData_2209_temp$Hour <= 18), ] #ok
MeasurementData_2209_temp<-MeasurementData_2209_temp[which(MeasurementData_2209_temp$Weekday != "Saturday" & MeasurementData_2209_temp$Weekday != "Sunday"), ]
```
Aggregating the data to be mean CO2 per hour
```{r}
aggred <- aggregate(MeasurementData_2209_co2[4], list(Room=MeasurementData_2209_co2$Room,Hour=MeasurementData_2209_co2$Hour), mean)
```
Same for temperature
```{r}
aggredtemp <- aggregate(MeasurementData_2209_temp[4], list(Room=MeasurementData_2209_temp$Room,Hour=MeasurementData_2209_temp$Hour), mean)
```
Plotting the dataset
```{r}
aggred %>%
  ggplot( aes(x=Hour, y=Value, group=Room, color=Room)) +
    geom_line() + ylab("Mean CO2 value")+theme_minimal()
```
Temperature plot
```{r}
aggredtemp %>%
  ggplot( aes(x=Hour, y=Value, group=Room, color=Room)) +
    geom_line() + ylab("Mean temperature")+theme_minimal()
```
Drop NAs to perform clustering
```{r}
aggred <- aggred %>%
  select(Room, Hour, Value) %>%
  drop_na() %>% # must drop NAs for clustering to work
  glimpse()
```
Create a data frame using the spread() function that lists every DataPointID with the mean CO2 value per hour
```{r}
spread_aggred <- aggred %>%
  spread(Room, Value)  %>%
  glimpse()
```
Run the hierarchical cluster analysis
Run the analysis by first transposing the dataframe into a matrix using t() and removing the hour variable using [-1] to remove the first row.

Next, calculate the Euclidean distance metric using the dist() function. Then use the hclust() function and the method=”ward.D” argument, to run the hierarchical cluster analysis.

```{r}
CO2 <- t(spread_aggred[-1])
CO2_dist <- dist(CO2, method="euclidean")  
fitCO2 <- hclust(CO2_dist, method="ward.D")  
```
Plot the clusters
Plot using base R’s plot() function and highlight a set number of clusters using rect.hclust() and the k= argument (number of clusters).
```{r}
plot(fitCO2)
rect.hclust(fitCO2, k=3, border="cadetblue")
```
Same diagram rotated
```{r}
library(ggdendro)
ggdendrogram(fitCO2, rotate = TRUE, theme_dendro = FALSE) +
  theme_minimal() + xlab("") + ylab("")
```
Merge the clusters into the full dataset
First, assign the four clusters to the data using cutree() and then, using tidy principles and after renaming columns and casting Rooms as characters, merge the clusters with the original dataset. 

```{r}
clustered_data <- cutree(fitCO2, k=3)
clustered_data_tidy <- as.data.frame(as.table(clustered_data)) %>% glimpse()
colnames(clustered_data_tidy) <- c("Room","cluster")
clustered_data_tidy$Room <- as.character(clustered_data_tidy$Room)
aggred$Room <- as.character(aggred$Room) #joining on char values
```
```{r}
joined_clusters <- aggred %>%
  inner_join(clustered_data_tidy, by = "Room") %>%
  glimpse()
```
Output joined clusters for further plotly charts
```{r}
write.csv(joined_clusters,"C:/Users/iced/Documents/HY/Year1/DS project/data/joined_clustersYSK.csv", row.names = FALSE)
```
How many Rooms per cluster:
```{r}
table(clustered_data_tidy$cluster)
```
Visualizing clusters
```{r}
cluster1 <- joined_clusters %>% filter(cluster == "1") 

cluster1 %>%
  ggplot( aes(x=Hour, y=Value, group=Room, color=Room)) +
    geom_line() + ylab("Mean CO2 value") + theme_minimal()
```
```{r}
cluster2 <- joined_clusters %>% filter(cluster == "2") 

cluster2 %>%
  ggplot( aes(x=Hour, y=Value, group=Room, color=Room)) +
    geom_line() + ylab("Mean CO2 value") + theme_minimal()
```

```{r}
cluster3 <- joined_clusters %>% filter(cluster == "3") 

cluster3 %>%
  ggplot( aes(x=Hour, y=Value, group=Room, color=Room)) +
    geom_line() + ylab("Mean CO2 value") + theme_minimal()
```


